---
title: "STAT 253 HW4"
author: "Emydius, Hari, Lucy"
date: "2022-11-1"
output: html_document
---

```{r hw2_setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```

### Research question {-} 

**Data context:**

“Music Dataset: 1950 to 2019,” collected in a research study in 2020 using Spotify’s Python package and other web APIs, is a dataset consisting of basic information and metadata for 28,372 lyrical tracks released from 1950 to 2019. Each row in the dataset represents a track with its name, artist, release year, genre, lyrics, main topic, and normalized metadata. Here, metadata refers to musical features of a track, such as: danceability, loudness, acousticness, instrumentalness, valence, and energy. Besides, the authors also include normalized data for each of the sixteen topics a track can demonstrate, which ultimately determines its main topic. Examples of such topics are: dating, violence, sadness, romantic, etc.

> The dataset contains some of the most popular music genres: blues, country, hiphop, jazz, pop, reggae, and rock. Songs of these genres can be easily identified by ear, although what makes them distinctive from each other is often unexplained. Can we identify a song’s genre based on its musical features? Which musical feature(s) contribute the most to a song’s genre? Such answers could help us understand how the music industry, or more fundamentally, our human brain, classifies songs.

Try to implement at least 2 different classification methods to answer your research question.

KNN and Random Forest

Reflect on the information gained from these two methods and how you might justify this method to others.



Keep in mind that the final project will require you to complete the pieces below. Use this as a guide for your work but don’t try to accomplish everything for HW4:

### Data Exploration {-}

```{r}
# Library statements
library(dplyr);
library(readr);
library(broom);
library(ggplot2);
library(tidymodels);
tidymodels_prefer();

# Read in data
music <- read_csv("tcc_ceds_music.csv")
# Clean data
music_clean <- select(music, - "...1", -artist_name, -track_name) %>% mutate(genre = factor(genre))
head(music_clean)

set.seed(253)
```

```{r}
genre_count <- music_clean %>% group_by(genre) %>% summarise(n = n()) %>% mutate(freq=n/sum(n))
genre_count
```

There are 7 genres in our dataset: blues, country, hip hop, jazz, pop, reggae, and rock. Pop has the largest frequency among all 7 genres at 24.8%, and hip hop has the lowest frequency at 3.1%. We will be answering our research question using random forest and k nearest neighbors. Our predictors would be musical features: danceability, loudness, acousticness, instrumentalness ,valence and energy.

### Random Forest{-}

```{r}
# Model Specification
rf_spec <- rand_forest() %>%
  set_engine(engine = 'ranger') %>% 
  set_args(mtry = NULL, # size of random subset of variables; default is floor(sqrt(number of total predictors))
           trees = 1000, # Number of trees
           min_n = 2,
           probability = FALSE, # FALSE: get hard predictions (not needed for regression)
           importance = 'impurity') %>% # we'll come back to this at the end
  set_mode('classification') # change this for regression

# Recipe
data_rec <- recipe(genre ~ danceability+loudness+acousticness+instrumentalness+valence+energy, data = music_clean)

# Workflows
data_wf_mtry2 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 2)) %>%
  add_recipe(data_rec)

data_wf_mtry3 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 3)) %>%
  add_recipe(data_rec)

## Create workflows for mtry = 4, 5, and 6

data_wf_mtry4 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 4)) %>% 
  add_recipe(data_rec)

data_wf_mtry5 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 5)) %>% 
  add_recipe(data_rec)

data_wf_mtry6 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 6)) %>% 
  add_recipe(data_rec)

```

```{r}
# Fit Models
set.seed(123) # make sure to run this before each fit so that you have the same 1000 trees
data_fit_mtry2 <- fit(data_wf_mtry2, data = music_clean)

set.seed(123) # make sure to run this before each fit so that you have the same 1000 trees
data_fit_mtry3 <- fit(data_wf_mtry3, data = music_clean)

# Fit models for 12, 74, 147
set.seed(123) 
data_fit_mtry4 <- fit(data_wf_mtry4, data = music_clean)

set.seed(123)
data_fit_mtry5 <- fit(data_wf_mtry5, data = music_clean)

set.seed(123) 
data_fit_mtry6 <- fit(data_wf_mtry6, data = music_clean)
```

```{r}
# Custom Function to get OOB predictions, true observed outcomes and add a user-provided model label
rf_OOB_output <- function(fit_model, model_label, truth){
    tibble(
          .pred_class = fit_model %>% extract_fit_engine() %>% pluck('predictions'), #OOB predictions
          class = truth,
          label = model_label
      )
}

#check out the function output
rf_OOB_output(data_fit_mtry2,2, music_clean %>% pull(genre))

```
```{r}
data_rf_OOB_output <- bind_rows(
    rf_OOB_output(data_fit_mtry2,2, music_clean %>% pull(genre)),
    rf_OOB_output(data_fit_mtry3,3, music_clean %>% pull(genre)),
    rf_OOB_output(data_fit_mtry4,4, music_clean %>% pull(genre)),
    rf_OOB_output(data_fit_mtry5,5, music_clean %>% pull(genre)),
    rf_OOB_output(data_fit_mtry6,6, music_clean %>% pull(genre))
)


data_rf_OOB_output %>% 
    group_by(label) %>%
    accuracy(truth = class, estimate = .pred_class)
```

```{r}
labels <- data_rf_OOB_output %>% 
    select(label)
accuracys <- data_rf_OOB_output %>% 
    group_by(label) %>% 
    accuracy(truth = class,estimate= .pred_class) 
    

ggplot(accuracys,aes(x= label,y=.estimate))+geom_point()+geom_line()
```
```{r}
model_output <-data_fit_mtry2 %>% 
    extract_fit_engine() 

model_output %>% 
    vip(num_features = 30) + theme_classic() #based on impurity

model_output %>% vip::vi() %>% head()
```
```{r}
model_output2 <- data_wf_mtry2 %>% 
  update_model(rf_spec %>% set_args(importance = "permutation")) %>% #based on permutation
  fit(data = music_clean) %>% 
    extract_fit_engine() 

model_output2 %>% 
    vip(num_features = 30) + theme_classic()


model_output2 %>% vip::vi() %>% head()
```
```{r}
ggplot(music_clean, aes(x = genre, y = danceability)) +
    geom_violin() + theme_classic()
ggplot(music_clean, aes(x = genre, y = loudness)) +
    geom_violin() + theme_classic()
```

```{r}
# CV Folds
data_cv10 <- vfold_cv(music_clean, v = 10)

# Model Specification
knn_spec <- 
  nearest_neighbor() %>% # new type of model!
  set_args(neighbors = tune()) %>% # tuning parameter is neighbor; tuning spec
  set_engine(engine = 'kknn') %>% # new engine
  set_mode('classification') 

# Recipe with standardization (!)
data_rec <- recipe( genre ~ danceability+loudness+acousticness+instrumentalness+valence+energy , data = music_clean) %>%
    step_nzv(all_predictors())# removes variables with the same value

# Workflow (Recipe + Model)
knn_wf <- workflow() %>%
  add_model(knn_spec) %>% 
  add_recipe(data_rec)

# Tune model trying a variety of values for neighbors (using 10-fold CV)
penalty_grid <- grid_regular(
  neighbors(range = c(1, 50)), #  min and max of values for neighbors
  levels = 50) # number of neighbors values

knn_fit_cv <- tune_grid(knn_wf, # workflow
              resamples = data_cv10, #CV folds
              grid = penalty_grid, # grid specified above
              metrics = metric_set(accuracy))


```

Classification - Methods

Indicate at least 2 different methods used to answer your classification research question.

```{r}

```


Describe what you did to evaluate the models explored.

```{r}

```

Indicate how you estimated quantitative evaluation metrics.

```{r}

```


Describe the goals / purpose of the methods used in the overall context of your research investigations.

Classification - Results

Summarize your final model and justify your model choice (see below for ways to justify your choice).

Compare the different classification models tried in light of evaluation metrics, variable importance, and data context.
Display evaluation metrics for different models in a clean, organized way. This display should include both the estimated metric as well as its standard deviation. (This won’t be available from OOB error estimation. If using OOB, don’t worry about reporting the SD.)
Broadly summarize conclusions from looking at these evaluation metrics and their measures of uncertainty.
Classification - Conclusions - Interpret evaluation metric(s) for the final model in context. Does the model show an acceptable amount of error? - If using OOB error estimation, display the test (OOB) confusion matrix, and use it to interpret the strengths and weaknesses of the final model. - Summarization should show evidence of acknowledging the data context in thinking about the sensibility of these results.