# ---
title: "STAT 253 HW4"
author: "Emydius, Hari, Lucy"
date: "2022-11-1"
output: html_document
---

```{r hw4_setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```

# Homework 4 {-} 

## Research question {-} 

**Data context:**

“Music Dataset: 1950 to 2019,” collected in a research study in 2020 using Spotify’s Python package and other web APIs, is a dataset consisting of basic information and metadata for 28,372 lyrical tracks released from 1950 to 2019. Each row in the dataset represents a track with its name, artist, release year, genre, lyrics, main topic, and normalized metadata. Here, metadata refers to musical features of a track, such as: danceability, loudness, acousticness, instrumentalness, valence, and energy. Besides, the authors also include normalized data for each of the sixteen topics a track can demonstrate, which ultimately determines its main topic. Examples of such topics are: dating, violence, sadness, romantic, etc.

> The dataset contains some of the most popular music genres: blues, country, hiphop, jazz, pop, reggae, and rock. Songs of these genres can be easily identified by ear, although what makes them distinctive from each other is often unexplained. **Can we identify a song’s genre based on its musical features? Which musical feature(s) contribute the most to a song’s genre?** Such answers could help us understand how the music industry, or more fundamentally, our human brain, classifies songs.


## Data Exploration {-}

First, let's explore our data in order to shape our model choice.

```{r}
# Library statements
library(dplyr);
library(readr);
library(broom);
library(ggplot2);
library(tidymodels);
library(vip);
tidymodels_prefer();

# Read in data
music <- read_csv("tcc_ceds_music.csv")
# Clean data
music_clean <- select(music, genre, danceability, loudness, acousticness, instrumentalness, valence, energy) %>%
  mutate(genre = factor(genre))

head(music_clean)

set.seed(253)
```

```{r}
genre_count <- music_clean %>% group_by(genre) %>% 
  summarise(n = n()) %>% mutate(freq=n/sum(n)) %>% arrange(-freq)
genre_count
```

There are 7 genres in our dataset: blues, country, hip hop, jazz, pop, reggae, and rock. Pop has the largest frequency among all 7 genres at 24.8%; hip hop has the lowest frequency at 3.1%, and reggae has the second lowest frequency at 8.8%.

Because our outcome variable has 7 possible values, we cannot use logistic regression which is used for binary classification. Hence, we will be answering our research question using random forest and K-nearest neighbors. 

However, we note that since all our predictors - danceability, loudness, acousticness, instrumentalness, valence, and energy, are all numerical, it will be highly computationally intensive to perform random forest because there are so many possible splits to consider. And although KNN generally has low computational time because of its lazy learning approach, when paired with cross-validation and tune for the best K, it can be computationally heavy as well. 


## Methods {-}

### Random Forest {-}

For our models specification, we'll start with the random forests of 1000 trees, and our trees will be complex with min_n of 10 (so all the leaf nodes will have at most 10 cases). To make our trees diverse and avoid overfitting (when all trees are too similar), each tree will have different predictors to consider at each split. This number of predictors also identifies the different forests we will test out to determine which `mtry` is best for our overall accuracy. Since we only have 6 predictors, we will test out every possible `mtry` and see which one has the best overall accuracy.

```{r}
# Model Specification
rf_spec <- rand_forest() %>%
  set_engine(engine = 'ranger') %>% 
  set_args(mtry = NULL, # size of random subset of variables; default is floor(sqrt(number of total predictors))
           trees = 1000, # Number of trees
           min_n = 2,
           probability = FALSE, # FALSE: get hard predictions (not needed for regression)
           importance = 'impurity') %>% # we'll come back to this at the end
  set_mode('classification') 

# Recipe
data_rec <- recipe(genre ~ danceability+acousticness+loudness+instrumentalness+valence+energy, data = music_clean)

# Workflows
## Create workflows for mtry = 2, 3, 4, 5, and 6
data_wf_mtry1 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 1)) %>%
  add_recipe(data_rec)

data_wf_mtry2 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 2)) %>%
  add_recipe(data_rec)

data_wf_mtry3 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 3)) %>%
  add_recipe(data_rec)

data_wf_mtry4 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 4)) %>% 
  add_recipe(data_rec)

data_wf_mtry5 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 5)) %>% 
  add_recipe(data_rec)

data_wf_mtry6 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 6)) %>% 
  add_recipe(data_rec)
```

```{r}
# Fit Models
# Fit models for mtry = 2, 3, 4, 5, and 6
set.seed(253) # make sure to run this before each fit so that you have the same 1000 trees
data_fit_mtry1 <- fit(data_wf_mtry1, data = music_clean)

set.seed(253)
data_fit_mtry2 <- fit(data_wf_mtry2, data = music_clean)

set.seed(253)
data_fit_mtry3 <- fit(data_wf_mtry3, data = music_clean)

set.seed(253) 
data_fit_mtry4 <- fit(data_wf_mtry4, data = music_clean)

set.seed(253)
data_fit_mtry5 <- fit(data_wf_mtry5, data = music_clean)

set.seed(253) 
data_fit_mtry6 <- fit(data_wf_mtry6, data = music_clean)
```

```{r}
# Custom Function to get OOB predictions, true observed outcomes and add a user-provided model label
rf_OOB_output <- function(fit_model, model_label, truth){
    tibble(
          .pred_class = fit_model %>% extract_fit_engine() %>% pluck('predictions'), #OOB predictions
          class = truth,
          label = model_label
      )
}

#check out the function output
rf_OOB_output(data_fit_mtry1,1, music_clean %>% pull(genre))
```



```{r}
data_rf_OOB_output <- bind_rows(
    rf_OOB_output(data_fit_mtry1,1, music_clean %>% pull(genre)),
    rf_OOB_output(data_fit_mtry2,2, music_clean %>% pull(genre)),
    rf_OOB_output(data_fit_mtry3,3, music_clean %>% pull(genre)),
    rf_OOB_output(data_fit_mtry4,4, music_clean %>% pull(genre)),
    rf_OOB_output(data_fit_mtry5,5, music_clean %>% pull(genre)),
    rf_OOB_output(data_fit_mtry6,6, music_clean %>% pull(genre))
)


data_rf_OOB_output %>% 
    group_by(label) %>%
    accuracy(truth = class, estimate = .pred_class)
```

```{r}
labels <- data_rf_OOB_output %>% 
    select(label)
accuracys <- data_rf_OOB_output %>% 
    group_by(label) %>% 
    accuracy(truth = class,estimate= .pred_class) 

ggplot(accuracys,aes(x= label,y=.estimate)) +
  geom_point() +
  geom_line() +
  theme_classic()
```

As it turns out, `data_fit_mtry1` gives us the best overall accuracy estimate. This accuracy drops with a decreasing speed as we increase `mtry` to 2, 3, 4, and 5. Then, when `mtry`=6, we have a slightly better accuracy compared to `data_fit_mtry5`. 

Considering the bias-variance tradeoff, although `mtry`=1 leads to very high variance in the decisions among the trees, we know that in the end, when counting the majority votes from all trees in the forest, this model still wins as the one whose majority votes are closest to the truth. This may be due to the fact that `data_fit_mtry1` was able to include splits on "weaker" variables that normally would not "win" in any splits. Thus, `data_fit_mtry1` has more chance to explore the globally best combination of splits, rather than locally best ones that prioritize strong variables or variables that are more correlated with the previously chosen one. 

To know the reason why this may be the fact, we take a look at the variable importance:

```{r}
# Impurity approach
impurity_output1 <-data_fit_mtry1 %>% 
    extract_fit_engine() 

impurity_output1 %>% 
    vip(num_features = 30) + theme_classic() #based on impurity

impurity_output1 %>% vip::vi() %>% head()
impurity_output1 %>% vip::vi() %>% tail()
```


```{r}
# Permutation approach
permutation_output1 <- data_wf_mtry1 %>% 
  update_model(rf_spec %>% set_args(importance = "permutation")) %>% #based on permutation
  fit(data = music_clean) %>% 
    extract_fit_engine() 

permutation_output1 %>% 
    vip(num_features = 30) + theme_classic()


permutation_output1 %>% vip::vi() %>% head()
```

Both measurements agree on the the rank of `danceability` (most important) and `valence` (5th rank), but swap the 2nd vs. 3rd ranks and 4th vs. 6th ranks. Impurity ranks `acousticness` higher than `energy`, and `instrumentalness` higher than `loudness`.

However, we know that the impurity measure tends to favor predictors with a lot of unique values, because with many unique values, there are many more possible splits, thus making it easier to find an optimal split that makes a big improvement in purity. Therefore, we believe that the result from the permutation approach is more reliable. 

To find out whether this is actually the case, we will take a look at the number of unique values of these predictors: 

```{r}
# Count the number of unique values of each predictor
music_clean %>% select(energy, acousticness, instrumentalness, loudness) %>%
  sapply(function (x) n_distinct(x))
```

As it turns out, impurity may have actually favored predictors with more unique values - `acousticness` has ~3 times more unique values than `energy`, and `loudness` also has ~3 times more unique values than `instrumentalness`. Therefore, the variable importance rankings of permutation is more likely to be accurate.

We can also check the violin plots of all predictors to verify the result of permutation:

```{r}
# Violin plots of the predictors, in order of importance according to permutation measurement
ggplot(music_clean, aes(x = genre, y = danceability)) + 
    geom_violin() + theme_classic()
ggplot(music_clean, aes(x = genre, y = energy)) + 
    geom_violin() + theme_classic()
ggplot(music_clean, aes(x = genre, y = acousticness)) + 
    geom_violin() + theme_classic()

ggplot(music_clean, aes(x = genre, y = instrumentalness)) + 
    geom_violin() + theme_classic()
ggplot(music_clean, aes(x = genre, y = valence)) + 
    geom_violin() + theme_classic()
ggplot(music_clean, aes(x = genre, y = loudness)) + 
    geom_violin() + theme_classic()
```

The rankings make sense considering how each predictor can single out certain genres.

Next, we will use an alternative model type to answer our research questions.

### K-nearest Neighbors {-}

```{r}
# CV Folds
data_cv10 <- vfold_cv(music_clean, v = 10)

# Model Specification
knn_spec <- 
  nearest_neighbor() %>% # new type of model!
  set_args(neighbors = tune()) %>% # tuning parameter is neighbor; tuning spec
  set_engine(engine = 'kknn') %>% # new engine
  set_mode('classification') 

# Recipe with standardization (!)
data_rec <- recipe( genre ~ danceability+loudness+acousticness+instrumentalness+valence+energy , data = music_clean) %>%
    step_nzv(all_predictors())# removes variables with the same value

# Workflow (Recipe + Model)
knn_wf <- workflow() %>%
  add_model(knn_spec) %>% 
  add_recipe(data_rec)

# Tune model trying a variety of values for neighbors (using 10-fold CV)
penalty_grid <- grid_regular(
  neighbors(range = c(1, 50)), #  min and max of values for neighbors
  levels = 50) # number of neighbors values

knn_fit_cv <- tune_grid(knn_wf, # workflow
              resamples = data_cv10, #CV folds
              grid = penalty_grid, # grid specified above
              metrics = metric_set(accuracy))
```

```{r}
knn_fit_cv %>% autoplot()
```

```{r}
knn_fit_cv %>% show_best(metric = 'accuracy') # Show evaluation metrics for different values of neighbors, ordered

knn_fit_cv %>% 
  select_by_one_std_err(metric = 'accuracy', desc(neighbors))  # Choose neighbors value that leads to the highest neighbors within 1 se of the lowest CV MAE
```


Describe the goals / purpose of the methods used in the overall context of your research investigations.

## Classification - Results

Summarize your final model and justify your model choice (see below for ways to justify your choice).

Compare the different classification models tried in light of evaluation metrics, variable importance, and data context.
Display evaluation metrics for different models in a clean, organized way. This display should include both the estimated metric as well as its standard deviation. (This won’t be available from OOB error estimation. If using OOB, don’t worry about reporting the SD.)
Broadly summarize conclusions from looking at these evaluation metrics and their measures of uncertainty.

## Classification - Conclusions

Interpret evaluation metric(s) for the final model in context. Does the model show an acceptable amount of error? - If using OOB error estimation, display the test (OOB) confusion matrix, and use it to interpret the strengths and weaknesses of the final model. - Summarization should show evidence of acknowledging the data context in thinking about the sensibility of these results.

The evaluation metric we got for the final model was 40.624% accuracy with a random forest of mtry 2. This means that using our model, we were able to predict the correct genre given our metrics about 40.6% of the time. This isn't a great amount of accuracy, since it means 60% of the time we'll get something that is not correct. However, this model is accurate enough to be significant in predicting genre. Our non-information rate for our data was 24.82%, which we got from the largest category in our data set, being the pop genre. If our model had guessed every single song was pop, the accuracy would have been 24.82%, so the model has to exceed this in order for it to have any significance in classifying genre. Our accuracy, 40.6% is 1.64 times more accurate than the non-information rate meaning it's at least significantly more accurate in predicting genre than straight up guessing would be. Still, it doesn't exceed 50% meaning the model would be wrong more times than it is correct. According to our random forest model, the most important variable was danceability shown by both permutation and impurity results with our random forest model; showing dacneability with the highest importance metric.

```{r}
# Confusion Matrix
rf_OOB_output(data_fit_mtry2, 2, music_clean %>% pull(genre)) %>%
  conf_mat(truth = class, estimate = .pred_class)
```
Looking at the confusion matrix, we can see that for each song the model predicted the correct genre more times than all the other genres. Looking at actual numbers, interpretations are difficult to make given that the proportions for each genre were not equal and varied wildly. One observation we can make is that for when the model predicted reggae, it turned out to be hip hop and jazz an equal amount of time, despite hip hop having the lowest proportion of songs (904 songs vs jazz's 3845 songs) meaning it overpredicted reggae for hip hop in comparison to other genres. Otherwise, there is too much variance between proportions to safely make assumptions about the model's strengths and weaknesses based on the confusion matrix alone.